---
title: "life expectancy code"
output: html_document
---

# Loading packages in
Loaded the "life expectancy.csv" dataset and downloaded ggplot2 functions.
```{r setup, include=FALSE}
library(readr)
dataset <- read_csv("life expectancy.csv")
library(ggplot2)
```

# Modelling Health Expenditure % vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Health Expenditure %`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy vs Health Expenditure %",
    x = "Health Expenditure %",
    y = "Life Expectancy"
  ) +
# regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```

```{r}
model <- lm(`Life Expectancy World Bank` ~ `Health Expenditure %`, data = dataset)
summary(model)
```

# Modelling Undernourishment % vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Prevelance of Undernourishment`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Prevelance of Undernourishment %",
    x = "Prevelance of Undernourishment",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Prevelance of Undernourishment`, data = dataset)
summary(model)
```

# Modelling CO2 vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `CO2`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs CO2",
    x = "CO2",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```

```{r}
model <- lm(`Life Expectancy World Bank` ~ `CO2`, data = dataset)
summary(model)
```

# Modelling Education Expenditure % vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Education Expenditure %`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Education Expenditure %",
    x = "Education Expenditure %",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Education Expenditure %`, data = dataset)
summary(model)
```

# Modelling Unemployment level vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Unemployment`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Unemployment",
    x = "Unemployment",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Unemployment`, data = dataset)
summary(model)
```

# Modelling Corruption level vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Corruption`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Corruption",
    x = "Corruption",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Corruption`, data = dataset)
summary(model)
```

# Modelling Sas=nitation score vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Sanitation`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Sanitation",
    x = "Sanitation",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Sanitation`, data = dataset)
summary(model)
```

# Modelling Injuries vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Injuries`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Injuries",
    x = "Injuries",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Injuries`, data = dataset)
summary(model)
```

# Modelling Communicable vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Communicable`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs Communicable",
    x = "Communicable",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Communicable`, data = dataset)
summary(model)
```

# Modelling NonCommunicable vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `NonCommunicable`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs NonCommunicable",
    x = "NonCommunicable",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `NonCommunicable`, data = dataset)
summary(model)
```

# Modelling Year vs. Life Expectancy

```{r}
ggplot(dataset, aes(x = `Year`, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.01, height = 0.5) +
  labs(
    title = "Life Expectancy World Bank vs NonCommunicable",
    x = "Year",
    y = "Life Expectancy World Bank"
  ) +
  # regression line
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  theme_minimal()

```
```{r}
model <- lm(`Life Expectancy World Bank` ~ `Year`, data = dataset)
summary(model)
```

# Modelling Country Income Group vs. Life Expectancy

```{r}

dataset$IncomeNum <- as.numeric(factor(
  dataset$IncomeGroup,
  levels = c(
    "Low income",
    "Lower middle income",
    "Upper middle income",
    "High income"
  )
))
ggplot(dataset, aes(x = IncomeNum, y = `Life Expectancy World Bank`)) +
  geom_jitter(alpha = 0.05, width = 0.05, height = 0.5) +
  geom_smooth(method = "lm", se = TRUE, color = "blue", linewidth = 1.2) +
  scale_x_continuous(
    breaks = 1:4,
    labels = c(
      "Low income",
      "Lower middle income",
      "Upper middle income",
      "High income"
    )
  ) +
  labs(
    title = "Life Expectancy by Income Group (Linear Trend)",
    x = "Income Group",
    y = "Life Expectancy (World Bank)"
  ) +
  theme_minimal()

```

```{r}
model <- lm(`Life Expectancy World Bank` ~ `IncomeNum`, data = dataset)
summary(model)
```

